---
title: "Pima_analysis_25244295"
author: "sai satya vamsi krishna raju dandu_25244295"
date: "2026-03-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warnings=FALSE}
library(tidyverse)
library(mice)
library(VIM)
library(rsample)
library(dcurves)
library(pROC)
```


```{r}

pima <- read_csv("C:/Users/krish/Downloads/diabetes.csv")
head(pima)
```


```{r}

predictors <- names(pima)[-9]  ##until the outcome column
missing_pct <- round(colSums(is.na(pima[,predictors])) / nrow(pima) * 100, 1)
data.frame(Predictor = predictors, `% Missing` = missing_pct)
```
Since there are no missng values flagged as NA but they exist as 0, which should be treated to proceed:

```{r}
pima$Glucose[pima$Glucose == 0]             <- NA
pima$BloodPressure[pima$BloodPressure == 0] <- NA
pima$SkinThickness[pima$SkinThickness == 0] <- NA
pima$Insulin[pima$Insulin == 0]             <- NA
pima$BMI[pima$BMI == 0]   
```


```{r}
# Now re-run your code — you should see:
predictors <- names(pima)[-9]  ##until the outcome column
missing_pct <- round(colSums(is.na(pima[,predictors])) / nrow(pima) * 100, 1)
data.frame(Predictor = predictors, `% Missing` = missing_pct)

```
```{r}

```

From the above output, i see the insulin(48.7) and SkinThickness(29.6) has the highest proposition of missing data.0.7% percentage of glucose is also missng.

```{r}

md.pattern(pima, rotate.names = TRUE)


aggr(pima, sortVars = TRUE, sortCombs = TRUE,
     col = c("steelblue", "red"),
     numbers = TRUE, prop = TRUE)
```

#multi variable imputation:

```{r}
pima.imputed <- mice(pima, m=5, method='pmm', seed=123, print=FALSE)
pima_complete <- mice::complete(pima.imputed)
```

```{r}
sum(is.na(pima_complete))
table(pima_complete$Outcome)       

summary(pima$Insulin)           # original (with NAs)
summary(pima_complete$Insulin)
```
By looking at the results of 


```{r}
set.seed(123)
split <- initial_split(pima_complete, prop=0.8, strata=Outcome)
train <- training(split)
test  <- testing(split)
```

```{r}

log_pima <- glm(Outcome ~ ., data = train, family = binomial())
summary(log_pima)
```


```{r}

pred_prob <- predict(log_pima,newdata = test, type = "response")
##converting the probs:
pred_class <- ifelse(pred_prob > 0.5, 1, 0)

roc_obj <- roc(test$Outcome, pred_prob)
auc(roc_obj)

#brier score:

breier_score <- mean((pred_prob - test$Outcome)^2)
breier_score
```
```{r}
logit_pred <- log(pred_prob / (1 - pred_prob))
cal_mod    <- glm(test$Outcome ~ logit_pred, family=binomial)
coef(cal_mod)[2]  
```


```{r}
##adding predictions to the test data
test$pred_prob <- pred_prob

# Run DCA
dca_result <- dca(Outcome ~ pred_prob, data=test)
plot(dca_result, smooth=TRUE)
```

Decision curve analysis confirmed that the logistic regression model provides positive net benefit across all clinically relevant threshold probabilities (5%–85%), outperforming both treat-all and treat-none strategies. This supports the model's clinical utility for diabetes screening regardless of the chosen decision threshold.


##to look at other models like random forest:
```{r}

set.seed(123)
train$Outcome <- factor(train$Outcome)

rf_mod  <- randomForest(Outcome ~ ., data=train, ntree=500)
rf_prob <- predict(rf_mod, newdata=test, type="prob")[,2]


roc_rf  <- roc(test$Outcome, rf_prob)
auc(roc_rf)

mean((rf_prob - as.numeric(as.character(test$Outcome)))^2)

```


```{r}


library(caret)
set.seed(123)

ctrl <- trainControl(
  method          = "cv",
  number          = 10,
  classProbs      = TRUE,        # needed for AUC
  summaryFunction = twoClassSummary,
  savePredictions = TRUE
)



train_cv       <- train
train_cv$Outcome <- factor(train_cv$Outcome, 
                           levels=c(0,1), 
                           labels=c("No","Yes"))

rf_cv <- train(
  Outcome ~ .,
  data      = train_cv,
  method    = "rf",
  metric    = "ROC",            # optimise by AUC
  trControl = ctrl,
  ntree     = 500
)

# CV Results
rf_cv
rf_cv$results   # AUC across folds
```


```{r}
glm_cv <- train(
  Outcome ~ .,
  data      = train_cv,
  method    = "glm",
  family    = "binomial",
  metric    = "ROC",
  trControl = ctrl
)


test_cv        <- test
test_cv$Outcome <- factor(test_cv$Outcome, 
                          levels=c(0,1), 
                          labels=c("No","Yes"))

# Logistic
glm_prob <- predict(glm_cv, newdata=test_cv, type="prob")[,"Yes"]

auc(roc(test$Outcome, glm_prob))
```

```{r}
```

